EGGs are adhesive!\\ An algebraic account of equality saturation

### 2025-09-25 jlamp/jlamp.tex
Update jlamp.tex

For the sake of readability, all the proofs appear in the appendices. 
This work is an extended version of~\cite{uovadesive}.
With respect to the conference paper, the way term graphs
are described is completely changed: they are now presented 
as an instance of the newly introduced category of 
left-monogamous hypergraphs. Also the notion of separated 
hypergraphs is original, so that Sections 3.1.1 and 3.2.1 are largely new,
and consequently so Sections 4.1.1, 4.2, and 5. 
%
Finally, also Section~6 on the use of DPO rewriting has been rewritten 
and expanded, trying to highlight the nuances concerning its application.
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.

@@

The fact that the left-hand side of a rule is the identity allows for the repeated application of such rule, hence the 
possibility to keep on performing the same rewriting step.
This is forbidden by using rules \emph{with negative application conditions}, given as the usual span 
$L \leftarrow K \rightarrow R$ plus an additional arrow $n: L\rightarrow N$, such that a match $m: L \to G$ is admissible if it cannot
be factorised through $n$~\cite{HabelHT96}. For e-term graphs and rules $L \leftarrow L \rightarrow R$, it suffices to choose $n$ 
as the right-hand side itself.
Indeed, consider again the rule $x / x \to 1$: once $1$ has been added once to the equivalence class of $x / x$, it is not necessary to
keep on adding it, since e.g. the two pictures below represent the same EGG
\[xxx\]
%
The theory of $\mathcal{M}$-adhesivity carries on for rules with negative application conditions, see~\cite{ehrig2012,ehrig2014adhesive}.

>>>>

For the sake of readability, all the proofs appear in the appendices. 
This work is an extended version of~\cite{uovadesive}.
Its overall structure reflects the one of the conference paper, 
but the way term graphs
are described is completely changed: They are now presented 
as an instance of the newly introduced category of 
left-monogamous hypergraphs. Also the notion of separated 
hypergraphs is original, so that Sections 3.1.1 and 3.2.1 are largely new,
and consequently so Sections 4.1.1, 4.2, and 5. 
%
Finally, also Section~6 on the use of DPO rewriting has been rewritten 
and expanded, trying to highlight the nuances concerning its application.
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.

@@

The fact that the left-hand side of a rule is the identity allows for the repeated application of such rule, hence the 
possibility to keep on performing the same rewriting step.
This is forbidden by using rules \emph{with negative application conditions}, given as the usual span 
$L \leftarrow K \rightarrow R$ plus an additional arrow $n: L\rightarrow N$, such that a match $m: L \to G$ is admissible if it cannot
be factorised through $n$~\cite{HabelHT96}. For e-term graphs and rules $L \leftarrow L \rightarrow R$, it suffices to choose $n$ 
as the right-hand side itself.
Indeed, consider again the rule $x / x \to 1$: once $1$ has been added once to the equivalence class of $x / x$, it is not necessary to
keep on adding it, since e.g. the two pictures below intuitively represent the same EGG
\[xxx\]
%
The theory of $\mathcal{M}$-adhesivity carries on for rules with negative application conditions, see~\cite{ehrig2012,ehrig2014adhesive}.
